Interpretable Machine Learning

### 2017-11-27 talk.Rmd
improve narrative
improving talk

@@

```{r load-data, warning=FALSE, include=FALSE}
library('mlr')
library('ggplot2')
source('code/prepare-kaggle-data.R')
```

@@

```{r, echo=FALSE}
knitr::include_graphics("images/done-here.gif")
```

@@

![](images/Hide-the-pain-harold-phone.jpg)

## "There is a problem with the model!"
![](images/age.jpeg)

## Individual Conditional Expectation

```{r}
ice = generatePartialDependenceData(mod, task, features =c('Age'), 
                                     individual = TRUE)
plotPartialDependence(ice) + scale_y_continuous(limits=c(0, NA))
```

## ICE, centered

```{r}
ice.centered = generatePartialDependenceData(mod, task, 
                features =c('Age'), individual = TRUE, center = list(Age=20))
plotPartialDependence(ice.centered)
```

## Partial dependence plots
```{r}
pdp = generatePartialDependenceData(mod, task, features =c('Age'))
plotPartialDependence(pdp) + scale_y_continuous(limits=c(0, NA))
```

@@

## Permutation feature importance

```{r, warning=FALSE}
feat.imp = getFeatureImportance(mod, type=1)
knitr::kable(t(feat.imp$res))
```

## Gender?

@@

![](images/hidden-pain-bias.jpg)

TODO: Add zoomed in version of Hidden Pain Harolds face

@@

## Shapley value

Based on Game theory. 
TODO: Add visualisation

## Shapley value

TODO: Calculate shapley value for one instance.

@@

## When is it important
http://people.dbmi.columbia.edu/noemie/papers/15kdd.pdf
Asthma example

## Different entry points
- Understand data better $\rightarrow$ do it!
- Use only interpretable models (sparse linear models, short decision trees, rules, ...) $\rightarrow$ does not always work. 
- **Apply model-agnostic methods to make black boxes interpretable** $\rightarrow$ cool, tell me more

## Model-agnostic methods
- Not bound to model
- Use always the same type of explanation
- ...
- Disadvantage: Can't take advantage of internal structures

@@

Read my book: https://christophm.github.io/xai-book/

>>>>

## 
Scenario: A company approaches you to predict data scientist salaries with machine learning.

@@

```{r load-data, warning=FALSE, include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
source('code/prepare-kaggle-data.R')
```

@@

```{r, echo=FALSE, out.width='70%', fig.align='center'}
knitr::include_graphics("images/done-here.gif")
```

## "There is a problem with the model!"
```{r, echo=FALSE, out.width='100%', fig.align='center'}
knitr::include_graphics("images/Hide-the-pain-harold-phone.jpg")
```

## "The older the applicants, the higher the predicted salary, regardless of skills."
```{r, echo=FALSE, out.width='100%', fig.align='center'}
knitr::include_graphics("images/age.jpeg")
```

# Individual Conditional Expectation

@@

```{r}
ice = generatePartialDependenceData(mod, task, features ='Age', 
                                    individual = TRUE)
plotPartialDependence(ice) + scale_y_continuous(limits=c(0, NA))
```

```{r}
ice.c = generatePartialDependenceData(mod, task, features ='Age', 
          individual = TRUE, center = list(Age=20))
plotPartialDependence(ice.c)
```

# Partial dependence plots

```{r}
pdp = generatePartialDependenceData(mod, task, features =c('Age'))
plotPartialDependence(pdp) + scale_y_continuous(limits=c(0, NA))
```

@@

# Permutation feature importance

```{r, warning=FALSE}
feat.imp = getFeatureImportance(mod, type=1)$res
dat = gather(feat.imp, key='Feature', value='Importance') %>% arrange(Importance)
dat$Feature = factor(dat$Feature, levels = dat$Feature)
ggplot(dat)  + geom_point(aes(y=Feature, x = Importance))
```

## Gender?!
```{r, echo=FALSE, fig.align='center', out.width='60%'}
knitr::include_graphics("images/big-mistake.png")
```

@@

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics("images/hidden-pain-bias.jpg")
```

@@

# LIME

## 
```{r, echo=FALSE, fig.align='center'}
set.seed(44)
```
```{r, echo=TRUE, fig.align='center'}
dat = getTaskData(task)
explanation <- lime(dat, mod)
# Explain new instance
explainer <- lime::explain(dat[3, ], explanation, n_features = 3)
plot_features(explainer, ncol=1)
```

@@

## When is it important
Interpretability of machine learning models is when the problem is fundamentally underspecified.

@@

Read my book about "Interpretable Machine Learning"
https://christophm.github.io/xai-book/
